Evaluating the Importance of Initial Conditions for Antarctic Sea Ice Seasonal Predictability with a Fully Coupled Model

### 2025-10-09 analysis/paper/access-ice.qmd
Final polish after printing

---
title: Evaluating Seasonal Predictability of Antarctic Sea Ice and the importance of initial conditions with fully coupled models
format:
  nature-pdf:
    # journal.cite-style is included in the tex file but ignored by pandoc if 
    # cite-method is not `natbib`.
    journal:
      cite-style: sn-mathphys-num
    # `citeproc` is the pandoc default. Set `cite-method: natbib` if required
    # to use the bst styles from the upstream template.
    cite-method: natbib
    keep-tex: true
    equal-margins: false
    classoption: referee
    fig-width: 5  
    fig-pos: "!htb"
header-includes: | 
   \usepackage[section]{placeins}
author:
  - name: Elio Campitelli
    affiliations:
      - name: Monash University, Australia,
        id: 1
        department: School of Earth, Atmosphere and Environment
        # address: 1737 Cambridge Street
        # city: Cambridge
        # state: MA
        # postal-code: 02138
      - name: Monash University, Australia
        id: 2
        department: Securing Antarctica’s Environmental Future
        # address: 1 Oxford Street
        # city: Cambridge
        # state: MA
        # postal-code: 02138
    email: elio.campitelli@monash.edu
    attributes:
      corresponding: true
  - name: Ariaan Purich
    affiliations:
      - ref: 1
      - ref: 2
  - name: Julie Arblaster
    affiliations:
      - ref: 1
      - ref: 2
  - name: Eun-Pa Lim
    affiliations:
      - department: Research, Bureau of Meteorology
        id: 3
  - name: Matthew Wheeler
    affiliations:
      - ref: 3      
  - name: Phillip Reid
    affiliations:
      - ref: 3      
keywords: [sea ice, predictabilty, initial conditions, forecasting]
bibliography: references.bib

@@

Accurate Antarctic sea-ice forecasts are crucial for climate monitoring and operational planning, yet they remain challenging due to model biases and complex ice-ocean-atmosphere interactions. 
We investigate the role of initial conditions on Antarctic sea-ice forecasts by comparing hindcasts from two versions of the Australian Bureau of Meteorology's ACCESS seasonal forecast system: +S1 and +S2. 
They use identical model configuration and differ only in their initial conditions; primarily in that +S2 does not assimilate sea-ice observations. 
Our analysis reveals that both systems experience an extended melt season and delayed growth phase. 
This leads to a significant negative sea-ice extent bias which is corrected only in +S1 by the data assimilation system. 
The impact of the initial conditions on forecast errors varies dramatically by season: summer and autumn initial conditions (January-April) provide predictive skill for up to three months, with February initial conditions being particularly crucial. 
In contrast, winter forecasts off the two systems are statistically indistinguishable after only two weeks. 
Regional analysis of forecast skill suggests that this winter predictability barrier is most dramatic over Eastern Antarctica, where evern +S1 shows negative skill. 
These findings highlight the critical importance of comprehensive year-round sampling in predictability studies and suggest that operational sea-ice data assimilation efforts should prioritize the summer-autumn period when initial conditions have maximum impact on forecast skill.

@@

However, producing accurate Antarctic sea-ice forecasts has been challenging due to model biases, inherent large variability and complexity, and they have lagged behind Arctic sea-ice forecasts[@zampieri2019; @gao2024a].
Dynamical seasonal forecasts of summer Antarctic sea ice have been shown to perform worse than relatively simple statistical methods [@massonnet2023] and machine learning approaches [e.g. @dong2024; @lin2025], which also underscores the need for better understanding of sea-ice dynamics, and drivers of its variability.

Good initial conditions are generally required for a good forecast, however, it is not entirely known to what extent accurate sea-ice initial conditions affect the quality of the forecast and at what timescales.
Exploring seasonal predictions of Arctic sea-ice, @guemas2016 found that sea-ice initial conditions are important in autumn to predict summer sea-ice, but the impact wasn’t as dramatic when predicting winter sea-ice.
@day2014 also found seasonally-varying differences in the effect of initialisation, noting that accurate Arctic sea-ice thickness lead to improved sea-ice forecasts initialised in July but not when initialised in January.

For the Antarctic, @holland2013 studied the initial-value predictability of Antarctic sea-ice in a perfect model study using the CCSM3 model.
They found that sea-ice and ocean initial conditions provide predictive information to forecast sea-ice edge location several months in advance and that some predictability is retained for up to two years thanks to ocean heat content anomalies that are advected eastward.
This is in contrast with @marchi2020, who ran perfect model experiments to argue that uncertainty in the predicted atmospheric state and evolution is the main driver of uncertainty in Antarctic sea-ice extent prediction on seasonal timescales, with sea-ice and ocean initial conditions having lesser importance.
More recently, @morioka2022 studied decadal forecasts of Antarctic sea-ice and found that initialising ocean and sea-ice improved the correlation between simulated and observed sea-ice concentration evolution in the Amundsen–Bellingshausen Sea.
It is hard to compare these studies since they are based on forecasts initialised at different times of the year and different frameworks: @holland2013 ran 20 ensemble members initialised on the 1st of January of a particular year, @marchi2020 ran forecasts from the 1st of March and 1st of September and @morioka2022 run forecasts only from the 1st of March.
@marchi2020 also used a coupled ocean–sea-ice model instead of a fully coupled model like @holland2013 did.
@morioka2022 used observed sea ice initial conditions and compared with observations, while @marchi2020 and @holland2013 are perfect model studies.

In October 2021 the Australian Bureau of Meteorology (BoM) upgraded the Australian Community Climate and Earth System Simulator – Seasonal (ACCES-S) from version S1 to S2.
While the base model remained the same, the change in version was focused on using ocean, sea-ice and land initial conditions generated by the BoM instead of depending on the UK Met Office.
Crucially, compared to +S1, +S2 does not assimilate sea-ice observations, so sea ice is only affected by the ocean and atmospheric data assimilation via the coupled integration.

Since model configuration is identical between +S1 and +S2, they form a sort of “natural experiment” where the same forecasting model was run over a long period of time with multiple ensemble forecasts initialised throughout the year, with the only difference being the initial conditions.
This provides an opportunity to test the effect of sea-ice initial conditions on the forecast of sea-ice concentrations and the climate.

In this study we compare sea-ice hindcasts produced by +S1 and +S2.
We focus on seasonality of errors and biases and the effect of the data assimilation system.
This comparison will also inform future work with the prediction system as a research tool to better understand the dynamics and variability of the Antarctic sea ice and its impacts on the climate system as well as to explore the potential of using its sea-ice forecasts for decision-making.
The work will also serve as a benchmark for future prediction systems to attempt to improve upon.

@@

Both systems take atmospheric initial conditions derived from ERA-interim [@dee2011] for their hindcasts.
The main difference between the hindcasts of the two systems are the ocean and sea-ice initial conditions.
+S1 ocean initial conditions come from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) data assimilation scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and sea-ice concentration.
+S2, on the other hand, is initialised from ocean conditions generated by the BoM weakly coupled ensemble data assimilation scheme described in @wedd2022.
This scheme uses an optimal interpolation method and assimilates temperature and salinity profiles from EN4 [@good2013].
SSTs are nudged to Reynolds OISSTv2.1 [@reynolds2007] in areas where SSTs are over 0ºC and Sea Surface Salinity is weakly nudged to the World Ocean Atlas 2013 climatology [@zweng2013].

Of most relevance for this work, sea-ice concentrations are not assimilated in +S2.
Assimilation cycles are performed daily.
The coupled model runs for 24 hours initialised from the previous cycle.
Then the restart file fields of the ocean component are used as first guess in the data assimilation cycle and the innovations are used to build the next ocean initial conditions for the following cycle.
The atmosphere fields from that daily integration are not used and instead the model atmosphere is initialised using ERA-Interim.
The sea-ice initial conditions for the next cycle are the unaltered output of the previous daily integration.
Then the cycle starts again and the coupled model runs for another 24 hours.
During this integration the sea-ice component is affected by the ocean innovations and the new atmosphere initial conditions via the coupler.

The +S2 hindcast set used in this study runs for the period 1981–2018. Each hindcast consists of nine ensemble members built from three consecutive 3-member forecasts initialised at the first of every month and the two previous days and run for 279 days. 
The +S1 hindcast set is built by perturbing the atmospheric fields only with a random field perturbation [@hudson2017] and runs for 217 days for the period 1990–2012.
We analise the ensemble mean hindcasts unless otherwise specified.

Anomalies for each hindcast set are taken with respect to their own climatology specific to each initialisation date and forecast lead time, for the period 1990–2012.
This serves as a first-order correction of model bias and drift.

@@

We compute RMSE as the square root of the area-averaged squared differences between grid-point forecasted and observed sea-ice concentration anomalies.
We compute a pan-Antarctic RMSE by averaging over the whole NOAA/NSIDC CDRV4 Southern Hemisphere domain, and also a zonally-varying RMSE computed over twenty-four 15° longitude slices around Antarctica.

@@

@fig-hindcast-extent shows the mean sea-ice extent of the +S2 and +S1 hindcasts (row a) and their differences from the mean sea-ice extent of NOAA/NSIDC CDRV4 (row b).
Mean extent at the first of every month is indicated with circles for the initial conditions and with triangles for the longest lead time possible for each model (between `r range_max_lag[model == "S2", min]` and `r range_max_lag[model == "S2", max]` days for +S2 and between `r range_max_lag[model == "S1", min]` and `r range_max_lag[model == "S1", max]` days for +S1).
At this long lead time, the information of the initial conditions is essentially lost and the forecast reverts to each model's preferred equilibrium state.

+S2 initial conditions (circles; @fig-hindcast-extent column 1) show an overall negative bias, especially in the late summer-early autumn, while +S1 initial conditions (circles in @fig-hindcast-extent column 2) are very close to observations, as expected from the assimilation of sea-ice observations to produce the initial conditions of +S1.
Both systems' equilibrium states (triangles) show negative biases of sea-ice extent, particularly in the growth phase of late-autumn and winter months.
This is due primarily to the melt season being longer than in observations and with faster melt between January and March and the growing seasons being shorter with slower growth during March and April.
This is then followed by faster growth between May and July (@fig-mean-growth).
Many sea-ice models exhibit this systematic underestimation during the sea-ice minimum and early freezing season [@massonnet2023], which could indicate problems in the representation of thermodynamics in the model [@zampieri2019].
It is also not surprising that both forecasting systems converge to a similar equilibrium state because they share the same model formulation.

The difference between the initial conditions (circles) and the model equilibrium state (triangles) can be mostly attributed to the effect of data assimilation, which in +S2 is due solely to atmospheric and oceanic data assimilation.
From May to October, in +S2 circles are closer to observations than to the triangles, indicating that the information from the ocean and atmosphere data assimilation is affecting sea ice and improving the initial conditions.
During these months, +S1 can overestimate the sea-ice extent at short lead time.
For the rest of the year circles are overlaid with triangles in +S2, indicating that the ocean and atmosphere data assimilation is not affecting sea ice and that this component of the model is virtually free-running.

@@

```{r fig-mean-growth}
#| fig-cap:  Mean daily sea-ice extent growth (10^6 km^2/day) in +S1 (green) and +S2 (purple) hindcasts and observations (black), computed as the mean daily differences in sea-ice extent between each date and the next for each forecast month. Values are smoothed with a 11-day running mean.
#| fig-height: 4

@@

+S1 has a comparatively smaller overall bias (@fig-bias-2). 
The largest values are found between April and June, when the faster growth resutls in large positive bias along the sea-ice edge, and in January, when the faster melt leads to large negative bias in the Weddell and Admunsen seas.

@@

hindast_extent_monthly |>
  _[, merge(.SD, monthly_extent, all = TRUE), by = .(model, lag)] |>
  _[lag %in% (c(1, 3, 5, 7) - 1)] |>
  _[time %between% s2_range] |>
  ggplot(aes(time)) +
  geom_line(aes(y = aice, group = dataset, color = dataset)) +
  geom_line(aes(y = forecast, color = model)) +
  geom_text(
    data = \(x) {
      x[
        time %between% s1_range,
        rmse(aice, forecast, signif = 2, scale = 1e12),
        by = .(lag, model)
      ]
    },
    aes(label = paste0("RMSE: ", text)),
    x = -Inf,
    y = Inf,
    size = 3,
    vjust = 1.2,
    hjust = 0
  ) +

  geom_text(
    data = \(x) {
      x[time %between% s1_range, correlate(aice, forecast), by = .(lag, model)]
    },
    aes(label = paste0("cor: ", text)),
    vjust = -0.8,
    hjust = 0,
    size = 3,
    x = -Inf,
    y = -Inf
  ) +

@@

Unexpectedly, for +S2, RMSE improves with lead time, even though the correlation degrades with lead time.
This is puzzling behaviour that goes contrary to what is usually seen in prediction models.
The explanation seems to be the mentioned increased interannual variability.
@fig-extent-sd shows monthly mean sea-ice extent of the forecasts as a function of lead time compared with observations.
+S1 standard deviation lies within the observed standard deviation regardless of lead time, while +S2 standard deviation is more than twice that of observations at zero lead time and only approaches the observed value at 9 month lead time for most months.

@@

  errors2 |>
    copy() |>
    _[obs_dataset == which_dataset] |>
    _[,
      base := low[version == "climatology"],
      by = .(lag, measure, month, obs_dataset)
    ] |>
    _[version != "climatology"] |>
    _[version != "persistence"] |>
    # _[lag > 2] |>
    _[, rle(estimate < base), by = .(version, measure, month, obs_dataset)] |>
    _[, .SD[1], by = .(version, measure, month, obs_dataset)] |>
    tidyr::complete(
      version,
      measure,
      month,
      obs_dataset,
      values,
      fill = list(lengths = 0)
    ) |>
    as.data.table() |>
    _[values == TRUE] |>
    _[, .(
      version,
      month,
      obs_dataset,
      lag = lengths,
      benchmark = "climatology"
    )]
) |>
  _[, time := make_date(2000, month, 1) + lag] |>
  # _[benchmark == "climatology" & version == "S1"] |>
  ggplot(aes(month, lag)) +
  ggbraid::geom_braid(
    data = \(x) dcast(x, month + lag + version ~ benchmark, value.var = "lag"),
    alpha = 0.3,
    aes(
      ymin = persistence,
      ymax = climatology,
      fill = climatology > persistence
    )
  ) +
  geom_line(aes(color = benchmark)) +
  # geom_point(aes(color = time), data = \(x) x[lag != 0]) +
  ggrepel::geom_text_repel(
    data = \(x) {
      x[, .SD[which.max(lag)], by = .(version, month, obs_dataset)] |>
        _[benchmark == "climatology"]
    },
    aes(label = format(time, "%d %b")),
    vjust = -0.4
  ) +
  geom_point(
    data = \(x) {
      x[, .SD[which.max(lag)], by = .(version, month, obs_dataset)] |>
        _[benchmark == "climatology"]
    },
    aes(color = benchmark)
  ) +
  scale_color_models +
  scale_y_continuous("Lead time (days)") +
  scale_x_continuous(
    "Forecast month",
    breaks = 1:12,
    labels = month.abb,
    minor_breaks = NULL
  ) +
  scale_fill_manual(
    guide = "none",
    values = c(`TRUE` = "seagreen", `FALSE` = "#FFFFFF00")
  ) +
  facet_wrap(~version, ncol = 1, labeller = labeller(version = labels_models)) +
  coord_cartesian(clip = "off") +
  tag_facets()

@@

To analyse the spatial distribution of the model error, we computed the RMSE of zonal mean sea-ice concentration anomalies on `r uniqueN(rmse_lon_mean$lon)` slices of `diff(unique(rmse_lon_mean$lon))[1]`° longitude span for each forecasting system.
We control for some areas being naturally easier to forecast than others by computing the RMSE skill score with the climatological forecast RMSE as reference.

@@

  variance_of_mean = clim_std2 |>
    rbind(errors) |>
    _[measure == "rmse"] |>
    _[lag %in% which_lags] |>
    _[measure == "rmse"] |>
    # _[version != "persistence"] |>
    _[obs_dataset == which_dataset] |>
    _[,
      .(value = var(value, na.rm = TRUE)),
      by = .(version, month(time_forecast), lag)
    ],
  idcol = "component"
) |>
  ggplot(aes(month, sqrt(value))) +
  geom_line(aes(color = version), linewidth = 1) +
  scale_y_continuous("Standard deviation") +
  # scale_color_models +
  scale_x_continuous("Month", breaks = seq(1, 12, by = 2), labels = \(x) {
    month.abb[x]
  }) +
  scale_color_models +
  facet_grid(
    lag ~ component,
    labeller = labeller(
      component = c(
        mean_variance = "Mean standard deviation\nof errors in ensemble members",
        variance_of_mean = "Standard deviation\nof ensemble mean error."
      )
    )
  ) +
  tag_facets(tag = "rc")

@@

In Figure 5 the mean error was shown. 
@fig-initial-spread column 1 shows the mean standard deviation of errors among ensemble members at various lead times.
At one day lead time ([Fig. @fig-initial-spread] a.1) +S2 has a slightly larger spread than +S1 due to the way that ensemble members are generated.
+S1 ensemble members are generated by adding random field perturbations to the atmosphere only, which then are transferred to the other components via the coupled simulation [@hudson2017]. 
With this scheme, ensemble members are all but guaranteed to be underdispersed in the ocean and sea-ice components. The time-lag ensemble used for +S2 ensures greater spread. 
This difference is gone after about just two days, and both systems have a comparable spread in ensemble member error afterwards ([Fig. @fig-initial-spread] b1 and c1).

@@

>>>>

---
title: Evaluating the Importance of Initial Conditions for Antarctic Sea Ice Seasonal Predictability with a Fully Coupled Model
format:
  nature-pdf:
    # journal.cite-style is included in the tex file but ignored by pandoc if 
    # cite-method is not `natbib`.
    journal:
      cite-style: sn-mathphys-num
    # `citeproc` is the pandoc default. Set `cite-method: natbib` if required
    # to use the bst styles from the upstream template.
    cite-method: natbib
    keep-tex: true
    equal-margins: false
    classoption: referee
    fig-width: 5  
    fig-pos: "!htb"
header-includes: | 
   \usepackage[section]{placeins}
author:
  - name: Elio Campitelli
    affiliations:
      - name: Monash University, Australia,
        id: 1
        department: School of Earth, Atmosphere and Environment
        # address: 1737 Cambridge Street
        # city: Cambridge
        # state: MA
        # postal-code: 02138
      - name: Monash University, Australia
        id: 2
        department: Securing Antarctica’s Environmental Future
        # address: 1 Oxford Street
        # city: Cambridge
        # state: MA
        # postal-code: 02138
    email: elio.campitelli@monash.edu
    attributes:
      corresponding: true
  - name: Ariaan Purich
    affiliations:
      - ref: 1
      - ref: 2
  - name: Julie Arblaster
    affiliations:
      - ref: 1
      - ref: 2
  - name: Eun-Pa Lim
    affiliations:
      - department: Research, Bureau of Meteorology
        id: 3
  - name: Matthew Wheeler
    affiliations:
      - ref: 3      
  - name: Phillip Reid
    affiliations:
      - ref: 3      
keywords: [sea ice, predictability, initial conditions, forecasting]
bibliography: references.bib

@@

Accurate Antarctic sea ice forecasts are crucial for climate monitoring and operational planning, yet they remain challenging due to model biases and complex ice-ocean-atmosphere interactions. 
The two versions of the Australian Bureau of Meteorology's ACCESS seasonal forecast system, +S1 and +S2, use identical model configuration and differ only in their initial conditions; primarily in that +S2 does not assimilate sea-ice observations. 
This makes for a good "natural experiment" to analyse the role of initial conditions on Antarctic sea ice forecasts using a long run and a moderately large ensemble with a fully coupled model.
Our analysis reveals that both systems experience an extended melt season and delayed growth phase. 
This leads to a significant negative sea-ice extent bias which is corrected only in +S1 by the data assimilation system. 
The impact of the initial conditions on forecast errors varies dramatically by season: summer and autumn initial conditions (January-April) provide predictive skill for up to three months, with February initial conditions being particularly crucial. 
In contrast, winter forecasts of the two systems are statistically indistinguishable after only two weeks. 
Regional analysis of forecast skill suggests that this winter predictability barrier is most dramatic over Eastern Antarctica, where even +S1 shows negative skill. 
These findings highlight the critical importance of comprehensive year-round sampling in predictability studies and suggest that operational sea-ice data assimilation efforts should prioritize the summer-autumn period when initial conditions have maximum impact on forecast skill.

@@

However, Antarctic sea ice forecasts have lagged behind Arctic sea ice forecasts due to model biases, inherent large variability and complexity [@zampieri2019; @gao2024a].
Dynamical seasonal forecasts of summer Antarctic sea ice have been shown to perform worse than relatively simpler statistical methods [@massonnet2023] and machine learning approaches (e.g. @dong2024, @lin2025), which also underscores the need for better understanding of sea-ice dynamics, and drivers of its variability.

Good initial conditions are generally required for a good forecast, however, it is not entirely known to what extent accurate sea ice initial conditions affect the quality of the forecast and at what timescales.
Exploring seasonal predictions of Arctic sea ice, @guemas2016 found that sea ice initial conditions are important in autumn to predict summer sea ice, but the impact wasn’t as dramatic when predicting winter sea ice.
@day2014 also found seasonally-varying differences in the effect of initialisation, noting that accurate Arctic sea-ice thickness lead to improved sea ice forecasts initialised in July but not when initialised in January.

For the Antarctic, @holland2013 studied the initial-value predictability of Antarctic sea ice in a perfect model study using the CCSM3 model.
They found that sea-ice and ocean initial conditions provide predictive information to forecast sea-ice edge location several months in advance and that some predictability is retained for up to two years thanks to ocean heat content anomalies that are advected eastward.
This is in contrast with @marchi2020, who ran perfect model experiments to argue that uncertainty in the predicted atmospheric state and evolution is the main driver of uncertainty in Antarctic sea-ice extent prediction on seasonal timescales, with sea-ice and ocean initial conditions having lesser importance.
More recently, @morioka2022 studied decadal forecasts of Antarctic sea ice and found that initialising ocean and sea ice improved the correlation between simulated and observed sea-ice concentration evolution in the Amundsen–Bellingshausen Sea.
It is hard to compare these studies since they are based on forecasts initialised at different times of the year and different frameworks: @holland2013 ran 20 ensemble members initialised on the 1st of January of a particular year, @marchi2020 ran forecasts from the 1st of March and 1st of September and @morioka2022 run forecasts only from the 1st of March.
@marchi2020 also used a coupled ocean–sea-ice model instead of a fully coupled model like @holland2013 did.
@morioka2022 used observed sea ice initial conditions and compared with observations, while @marchi2020 and @holland2013 are perfect model studies.

In October 2021 the Australian Bureau of Meteorology (BoM) upgraded the Australian Community Climate and Earth System Simulator – Seasonal (ACCESS-S) from version S1 to S2.
While the base model remained the same, the change in version was focused on using ocean, sea-ice and land initial conditions generated by the BoM instead of depending on the UK Met Office.
Crucially, compared to +S1, +S2 does not assimilate sea-ice observations, so sea ice is only affected by the ocean and atmospheric data assimilation via the coupled integration.

Since model configuration is identical between +S1 and +S2, they form a sort of “natural experiment” where the same forecasting model was run over a long period of time with multiple ensemble forecasts initialised throughout the year, with the only difference being the initial conditions.
This provides an opportunity to test the effect of sea ice initial conditions on the forecast of sea-ice concentrations and the climate.

In this study we compare sea ice hindcasts produced by +S1 and +S2.
We focus on seasonality of errors and biases and the effect of the data assimilation system.
This comparison will also inform future work with the prediction system as a research tool to better understand the dynamics and variability of the Antarctic sea ice and its impacts on the climate system as well as to explore the potential of using its sea ice forecasts for decision-making.
The work will also serve as a benchmark for future prediction systems to attempt to improve upon.

@@

Both systems take atmospheric initial conditions derived from ERA-interim [@dee2011] for their hindcasts.
The main difference between the hindcasts of the two systems are the ocean and sea ice initial conditions.
+S1's ocean and sea ice initial conditions come from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) data assimilation scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and satellite observations of sea-ice concentration using the EUMETSAT OSISAF product described in the next section.
+S2, on the other hand, is initialised from ocean conditions generated by the BoM weakly coupled ensemble data assimilation scheme described in @wedd2022.
This scheme uses an optimal interpolation method and assimilates temperature and salinity profiles from EN4 [@good2013].
SSTs are nudged to Reynolds OISSTv2.1 [@reynolds2007] in areas where SSTs are over 0ºC and Sea Surface Salinity is weakly nudged to the World Ocean Atlas 2013 climatology [@zweng2013].

Of most relevance for this work, sea-ice concentrations are not assimilated in +S2.
Assimilation cycles are performed daily.
The coupled model runs for 24 hours initialised from the previous cycle.
Then the restart file fields of the ocean component are used as first guess in the data assimilation cycle and the innovations are used to build the next ocean initial conditions for the following cycle.
The atmosphere fields from that daily integration are not used and instead the model atmosphere is initialised using ERA-Interim.
The sea ice initial conditions for the next cycle are the unaltered output of the previous daily integration.
Then the cycle starts again and the coupled model runs for another 24 hours.
During this integration the sea-ice component is affected by the ocean innovations and the new atmosphere initial conditions via the coupler.

The +S1 hindcast set is made up of nine members created by perturbing the atmospheric fields only with a random field perturbation [@hudson2017] and runs for 217 days for the period 1990–2012 initialised at the first of every month.
The +S2 hindcast set used in this study runs for the period 1981–2018. Each hindcast consists of nine ensemble members built from three consecutive 3-member forecasts created in the same manner as +S1 members initialised at the first of every month and the two previous days and run for 279 days. 
We analyse the ensemble mean hindcasts unless otherwise specified.

Anomalies for each hindcast set are taken with respect to their own climatology specific to each initialisation date and forecast lead time, for the period 1990–2012.
This serves as a first-order correction of model bias and drift.
For monthly means, we define "0 lead time months" as the monthly mean forecast of the same month of initialisation.

@@

We compute RMSE as the square root of the area-averaged squared differences between grid-point forecasted and observed sea-ice concentration anomalies.
We compute a pan-Antarctic RMSE by averaging over the whole NOAA/NSIDC CDRV4 Southern Hemisphere domain, and also a zonally-varying RMSE computed over 15 longitude slices 24° wide around Antarctica.

@@

@fig-hindcast-extent shows mean sea-ice extent of the +S1 and +S2 hindcasts (row a) and their differences from mean sea-ice extent of +CDR (row b).
Mean extent at the first of every month is indicated with circles for the initial conditions and with triangles for the longest lead time possible for each model (between `r range_max_lag[model == "S2", min]` and `r range_max_lag[model == "S2", max]` days for +S2 and between `r range_max_lag[model == "S1", min]` and `r range_max_lag[model == "S1", max]` days for +S1).
At this long lead time, information of the initial conditions is essentially lost and the forecast reverts to each model's preferred equilibrium state.

+S2 initial conditions (circles in [Fig. @fig-hindcast-extent] column 2) show an overall negative bias, especially in the late summer-early autumn, while +S1 initial conditions (circles in [Fig. @fig-hindcast-extent] column 1) are very close to observations, as expected from the assimilation of sea-ice observations to produce the initial conditions of +S1.
Both systems' equilibrium states (triangles) show negative biases of sea-ice extent, particularly in the growth phase of late-autumn and winter months.
This is due primarily to the melt season being longer than in observations and with faster melt between January and March and the growing seasons being shorter with slower growth during March and April.
This is then followed by faster growth between May and July (@fig-mean-growth).
Many sea-ice models exhibit this systematic underestimation during the sea-ice minimum and early freezing season [@massonnet2023], which could indicate problems in the representation of thermodynamics in the model [@zampieri2019].
It is also not surprising that both forecasting systems converge to a similar equilibrium state because they share the same model formulation.

The difference between the initial conditions (circles) and the model equilibrium state (triangles) can be mostly attributed to the effect of data assimilation, which in +S2 is due solely to the coupling of sea-ice with the atmosphere and the ocean. 
From May to October, in +S2 circles are closer to observations than the triangles are, indicating that the information from the ocean and atmosphere data assimilation is affecting sea ice and improving the initial conditions.
During these months, +S1 can overestimate the sea-ice extent at short lead time.
For the rest of the year circles are overlaid with triangles in +S2, indicating that the ocean and atmosphere data assimilation is not affecting sea ice and that this component of the model is virtually free-running.

@@

```{r fig-mean-growth}
#| fig-cap:  Mean daily sea-ice extent growth ($10^6 km^2/day$) in +S1 (green) and +S2 (purple) hindcasts and observations (black), computed as the mean daily differences in sea-ice extent between each date and the next for each forecast month. Values are smoothed with a 11-day running mean.
#| fig-height: 4

@@

+S1 has a comparatively smaller overall bias (@fig-bias-2). 
The largest values are found between April and June, when the faster growth results in large positive bias along the sea-ice edge, and in January, when the faster melt leads to large negative bias in the Weddell and Admunsen seas.

@@

hindast_extent_monthly |>
  _[, merge(.SD, monthly_extent, all = TRUE), by = .(model, lag)] |>
  _[lag %in% (c(1, 3, 5, 7) - 1)] |>
  _[time %between% s2_range] |>
  ggplot(aes(time)) +
  geom_line(aes(y = aice, group = dataset, color = dataset), linewidth = 0.2) +
  geom_line(aes(y = forecast, color = model), linewidth = 0.2) +
  geom_text(
    data = \(x) {
      x[
        time %between% s1_range,
        rmse(aice, forecast, signif = 2, scale = 1e12),
        by = .(lag, model)
      ]
    },
    aes(label = paste0("RMSE: ", text)),
    x = -Inf,
    y = Inf,
    size = 2.5,
    vjust = 2.5,
    hjust = 0
  ) +

  geom_text(
    data = \(x) {
      x[time %between% s1_range, correlate(aice, forecast), by = .(lag, model)]
    },
    aes(label = paste0("cor: ", text)),
    vjust = -1.5,
    hjust = 0,
    size = 2.5,
    x = -Inf,
    y = -Inf
  ) +

@@

Unexpectedly, for +S2, RMSE improves with lead time, even though the correlation degrades with lead time.
This is puzzling behaviour that goes contrary to what is usually seen in prediction models.
The explanation seems to be the mentioned increased interannual variability.
@fig-extent-sd shows the interannual standard deviation of monthly sea-ice extent of the forecasts as a function of lead time compared with observations.
+S1 standard deviation lies within the observed standard deviation regardless of lead time, while +S2 standard deviation is more than twice that of observations at zero lead time and only approaches the observed value at 9 month lead time for most months.

@@

  errors2 |>
    copy() |>
    _[obs_dataset == which_dataset] |>
    _[,
      base := low[version == "climatology"],
      by = .(lag, measure, month, obs_dataset)
    ] |>
    _[version != "climatology"] |>
    _[version != "persistence"] |>
    # _[lag > 2] |>
    _[, rle(estimate < base), by = .(version, measure, month, obs_dataset)] |>
    _[, .SD[1], by = .(version, measure, month, obs_dataset)] |>
    tidyr::complete(
      version,
      measure,
      month,
      obs_dataset,
      values,
      fill = list(lengths = 0)
    ) |>
    as.data.table() |>
    _[values == TRUE] |>
    _[, .(
      version,
      month,
      obs_dataset,
      lag = lengths,
      benchmark = "climatology"
    )]
) |>
  _[, time := make_date(2000, month, 1) + lag] |>
  # _[benchmark == "climatology" & version == "S1"] |>
  ggplot(aes(month, lag)) +
  ggbraid::geom_braid(
    data = \(x) dcast(x, month + lag + version ~ benchmark, value.var = "lag"),
    alpha = 0.3,
    aes(
      ymin = persistence,
      ymax = climatology,
      fill = climatology > persistence
    )
  ) +
  geom_line(aes(color = benchmark)) +
  # geom_point(aes(color = time), data = \(x) x[lag != 0]) +
  ggrepel::geom_text_repel(
    data = \(x) {
      x[, .SD[which.max(lag)], by = .(version, month, obs_dataset)] |>
        _[benchmark == "climatology"]
    },
    aes(label = format(time, "%d %b")),
    size = 2,
    vjust = -0.4
  ) +
  geom_point(
    data = \(x) {
      x[, .SD[which.max(lag)], by = .(version, month, obs_dataset)] |>
        _[benchmark == "climatology"]
    },
    aes(color = benchmark)
  ) +
  scale_color_models +
  scale_y_continuous("Lead time (days)") +
  scale_x_continuous(
    "Forecast month",
    breaks = 1:12,
    labels = month.abb,
    minor_breaks = NULL
  ) +
  scale_fill_manual(
    guide = "none",
    values = c(`TRUE` = "seagreen", `FALSE` = "#FFFFFF00")
  ) +
  facet_wrap(~version, ncol = 1, labeller = labeller(version = labels_models)) +
  coord_cartesian(clip = "off") +
  tag_facets()

@@

To analyse the spatial distribution of the model error, we computed the RMSE of zonal mean sea-ice concentration anomalies on `r uniqueN(rmse_lon_mean$lon)` slices of `r diff(unique(rmse_lon_mean$lon))[1]`° longitude span for each forecasting system.
We control for some areas being naturally easier to forecast than others by computing the RMSE skill score with the climatological forecast RMSE as reference.

@@

  variance_of_mean = clim_std2 |>
    rbind(errors) |>
    _[measure == "rmse"] |>
    _[lag %in% which_lags] |>
    _[measure == "rmse"] |>
    # _[version != "persistence"] |>
    _[obs_dataset == which_dataset] |>
    _[,
      .(value = var(value, na.rm = TRUE)),
      by = .(version, month(time_forecast), lag)
    ],
  idcol = "component"
) |>
  ggplot(aes(month, sqrt(value))) +
  geom_line(aes(color = version), linewidth = 0.5) +
  scale_y_continuous("Standard deviation") +
  # scale_color_models +
  scale_x_continuous("Month", breaks = seq(1, 12, by = 2), labels = \(x) {
    month.abb[x]
  }) +
  scale_color_models +
  facet_grid(
    lag ~ component,
    labeller = labeller(
      component = c(
        mean_variance = "Mean standard deviation\nof errors in ensemble members",
        variance_of_mean = "Standard deviation\nof ensemble mean error."
      )
    )
  ) +
  tag_facets(tag = "rc")

@@

In @fig-rmse the mean error was shown. 
@fig-initial-spread column 1 shows the mean standard deviation of errors among ensemble members at various lead times.
At one day lead time ([Fig. @fig-initial-spread] a.1) +S2 has a slightly larger spread than +S1 due to the way that ensemble members are generated.
+S1 ensemble members are generated by adding random field perturbations to the atmosphere only, which then are transferred to the other components via the coupled simulation [@hudson2017]. 
With this scheme, ensemble members are all but guaranteed to be underdispersed in the ocean and sea-ice components. The time-lag ensemble used for +S2 ensures greater spread. 
This difference is gone after about just two days, and both systems have a comparable spread in ensemble member error afterwards ([Fig. @fig-initial-spread] b1 and c1).

@@

\clearpage
